Global code motion of congruent instructions

### 2017-11-08 cgo/paper.tex
candidates to schedule

Dominance graphs, Global Value Numbers and Memory SSA are already available in
\LLVM{} but there is no facility to infer liveness of virtual registers. So we
implemented DJ-Graph data structure that allowed us to calculate MergeSets which
was used in the liveness analysis during \GCM{}.

@@

\section{A fast algorithm for code motion congruent computations}
The algorithm factors out the reachability of values such that multiple queries
to find reachability of values are fast. This is based on finding the dominance
frontiers and the post-dominance frontiers (PDF) of the CFG. These points, being
structural properties of a graph, do not change during code-motion. A
post-dominance frontier is the basic block where anticipability (ANTIC) of an
instruction (I) is checked because it the point on which the basic block
containing I is control dependent.

@@

A CHI keeps information about values flowing out of a basic block.  It is
similar to $\phi$ \cite{ssa} but in the inverse graph, and used for out-flowing
values on each edge. Although it may be suitable for other
analyses/optimizations, for \GCM{}, it is computed only for instructions with
multiple occurrences in the CFG. An illustration of CHI is shown in
Figure~\ref{fig:chi-intro}. The \GVN{} for both I1 and I2 is V, the CHI node
will save the instruction as well as the edge where the value is flowing to.
Where $Arg1$, $Arg2$ are of type CHIArg. The CHI for both I1 and I2 is
inserted in their PDF (A). For sinking we use the similar approach as
\cite{ssapre} to track values flowing into a basic block.

@@

When SSAPRE inserts $\Phi$ (expression PHIs), it does so at places where
(potentially) multiple expressions may merge. If a $\Phi$ has a missing entry,
that means the expression is partially available.  We apply a similar concept
for hoisting expression by working on inverted graph. We insert a CHI, to a
basic block with multiple successors. Then for each instruction, we register its
flow of value to its corresponding CHI which can be found in its PDF. For
conciseness we only register values with multiple occurrences. Finally, we start
walking the CFG and remove the CHI if it has a missing entry, that means the
expression is not anticipable in the basic block having CHI. For the remaining
CHIs anticipabilty of `values' is ensured by construction.

@@

\subsection{Finding candidates to move}
\label{subsec:finding-candidates}
The first step is to find a set of congruent instructions
\cite{briggs1997}. This is performed by a linear scan of all instructions of the
program and classifying them by their value numbers.

>>>>

Dominator tree, Global Value Numbers and Memory SSA are already available in
\LLVM{} but there is no facility to infer liveness of virtual registers. So we
implemented DJ-Graph data structure that allowed us to calculate MergeSets which
was used in the liveness analysis during \GCM{}.

@@

\section{A fast algorithm for code motion congruent computations}
The algorithm factors out the reachability of values such that multiple queries
to find reachability of values are fast. This is based on finding the dominance
frontiers and the post-dominance frontiers (PDF) of the CFG. These points, being
structural properties of a graph, do not change during code-motion. A
post-dominance frontier is the basic block where anticipability (ANTIC) of an
instruction (I) is checked because it is the basic block on which the basic
block containing I is control dependent.

@@

A CHI keeps information about `congruent' values flowing out of a basic block.
It is similar to $\phi$ \cite{ssa} but in the inverse graph, and used for
trackig out-flowing values on each edge. Although it may be suitable for other
analyses/optimizations, for \GCM{}, it is computed only for instructions with
multiple occurrences in the CFG. An illustration of CHI is shown in
Figure~\ref{fig:chi-intro}. The \GVN{} for both I1 and I2 is V, the CHI node
will save the instruction as well as the edge where the value is flowing to.
Where $Arg1$, $Arg2$ are of type CHIArg. The CHI for both I1 and I2 is inserted
in their PDF (A). For sinking we use the similar approach as \cite{ssapre} to
track values flowing into a basic block.

@@

When SSAPRE inserts $\Phi$ (expression PHIs), it does so at places where
(potentially) multiple expressions may merge. If a $\Phi$ has a missing entry,
that means the expression is partially available.  We apply a similar concept
for hoisting expression by working on inverted graph. We insert a CHI, to a
basic block with multiple successors. Then for each instruction, we register its
flow of value to its corresponding CHI which can be found in its PDF. For
conciseness we only register values with multiple occurrences. Finally, we start
walking the CFG and remove the CHI if it has a missing entry (see
Figure~\cite{fig:chi-intro2}), that means the expression is not anticipable in
the basic block having CHI. For the remaining CHIs anticipabilty of `values' is
ensured by construction.

@@

\subsection{Finding candidates to schedule}
\label{subsec:finding-candidates}
The first step is to find a set of congruent instructions
\cite{briggs1997}. This is performed by a linear scan of all instructions of the
program and classifying them by their value numbers.
